---
title: "Implementation of Adair & Lopez (2018) Method"
author: "Jeremy Roth"
date: "17 July 2020"
output:
  html_document:
    number_sections: yes
    toc: yes
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Setting up
## Load required packages
```{r, warning=FALSE, message=FALSE}
library(tidyverse)
```

## Specify directories where data and functions were saved
```{r, warning=FALSE, message=FALSE}
functions_dir <- "" # working directory 
data_dir <- "" # working directory 
```

## Load our functions to implement the Adair & Lopez (2018) method
```{r, warning=FALSE, message=FALSE}
source(paste0(functions_dir,
              "fn_Adair_Lopez.R"))
```

## Load coefficient estimates from Adair & Lopez (2018)
```{r, warning=FALSE, message=FALSE}
load(paste0(data_dir,
            "Adair_Lopez_Coefficient_Estimates.Rdata"))
```

## Load example dataset from supplemental material of Adair & Lopez (2018)
```{r, warning=FALSE, message=FALSE}
national_data <- read.csv(paste0(data_dir,
                                 "Supporting_information_data.csv"),
                          stringsAsFactors=FALSE)
# remove spaces from country names to help with merging in estimates of random intercepts
national_data$Country <- str_replace_all(string=national_data$Country,
                                         pattern=" ",
                                         replacement="")
# partition national-level dataset into three datasets: females, males, and both sexes combined
national_data_both <- national_data %>% filter(sex == 3) ## combined for both sexes
national_data_females <- national_data %>% filter(sex == 2) ## females only
national_data_males <- national_data %>% filter(sex == 1) ## males only
```

# Compute estimates with Model 1 (under-five completeness included as predictor)
## Model 1 - females and males combined
```{r, warning=FALSE, message=FALSE}
results_model1_both <- AdairLopezPrediction(data=national_data_both,
                                            model.type="model1_both",
                                            name.regCDR.per.1000="reg_cdr",
                                            name.pct65.as.decimal="pct65",
                                            name.under5.mortality.per.1000.livebirths="u5mr",
                                            name.year="year",
                                            name.country="Country",
                                            name.under5.completeness.as.decimal="u5compl")
results_model1_both %>% select(Country, year, sex, AL_predicted_completeness) %>%
                        head()
```

## Model 1 - females only
```{r, warning=FALSE, message=FALSE}
results_model1_females <- AdairLopezPrediction(data=national_data_females,
                                               model.type="model1_females",
                                               name.regCDR.per.1000="reg_cdr",
                                               name.pct65.as.decimal="pct65",
                                               name.under5.mortality.per.1000.livebirths="u5mr",
                                               name.year="year",
                                               name.country="Country",
                                               name.under5.completeness.as.decimal="u5compl")
results_model1_females %>% select(Country, year, sex, AL_predicted_completeness) %>%
                           head()
```

## Model 1 - males only
```{r, warning=FALSE, message=FALSE}
results_model1_males <- AdairLopezPrediction(data=national_data_males,
                                             model.type="model1_males",
                                             name.regCDR.per.1000="reg_cdr",
                                             name.pct65.as.decimal="pct65",
                                             name.under5.mortality.per.1000.livebirths="u5mr",
                                             name.year="year",
                                             name.country="Country",
                                             name.under5.completeness.as.decimal="u5compl")
results_model1_males %>% select(Country, year, sex, AL_predicted_completeness) %>%
                         head()
```


# Compute estimates with Model 2 (under-five completeness not included as predictor)
## Model 2 - females and males combined
```{r, warning=FALSE, message=FALSE}
results_model2_both <- AdairLopezPrediction(data=national_data_both,
                                            model.type="model2_both",
                                            name.regCDR.per.1000="reg_cdr",
                                            name.pct65.as.decimal="pct65",
                                            name.under5.mortality.per.1000.livebirths="u5mr",
                                            name.year="year",
                                            name.country="Country",
                                            name.under5.completeness.as.decimal="u5compl")
results_model2_both %>% select(Country, year, sex, AL_predicted_completeness) %>%
                        head()
```

## Model 2 - females only
```{r, warning=FALSE, message=FALSE}
results_model2_females <- AdairLopezPrediction(data=national_data_females,
                                               model.type="model2_females",
                                               name.regCDR.per.1000="reg_cdr",
                                               name.pct65.as.decimal="pct65",
                                               name.under5.mortality.per.1000.livebirths="u5mr",
                                               name.year="year",
                                               name.country="Country",
                                               name.under5.completeness.as.decimal="u5compl")
results_model2_females %>% select(Country, year, sex, AL_predicted_completeness) %>%
                           head()
```

## Model 2 - males only
```{r, warning=FALSE, message=FALSE}
results_model2_males <- AdairLopezPrediction(data=national_data_males,
                                             model.type="model2_males",
                                             name.regCDR.per.1000="reg_cdr",
                                             name.pct65.as.decimal="pct65",
                                             name.under5.mortality.per.1000.livebirths="u5mr",
                                             name.year="year",
                                             name.country="Country",
                                             name.under5.completeness.as.decimal="u5compl")
results_model2_males %>% select(Country, year, sex, AL_predicted_completeness) %>%
                         head()
```

# References
Adair, Tim, and Alan D. Lopez. "Estimating the completeness of death registration: an empirical method." PloS one 13.5 (2018): e0197047.
